---
title: "p8105_hw2_xf2226"
author: "Xin Fang"
date: "10/3/2021"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1 

## Read in and clean Mr. Trash Wheel sheet
```{r load and clean wheel data, message=F, warning=F}
mr_trash_wheel_df <- readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = cellranger::cell_cols("A:N"), skip = 1) %>% # specified the sheet in the Excel file, using sheet = 1; omited non-data entries, using range = cellranger::cell_cols("A:N") and skip = 1
  janitor::clean_names() %>% #use reasonable variable names
  filter(!is.na(dumpster)) #omit rows that do not include dumpster-specific data

mr_trash_wheel_df$sports_balls <- round(pull(mr_trash_wheel_df, sports_balls), 0) #round the number of sports balls to the nearest integer
```

## Read and clean precipitation data for 2018 and 2019

```{r Read and clean precipitation data for 2018 and 2019, message=F, warning=F}
sheets <- readxl::excel_sheets("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx") #read in sheets

precipitation_2018_2019 <- do.call(rbind, lapply(sheets[4:5], function(X) 
        transform(readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = X, skip = 1), sheetname = X))) #rbind the two seperate sheets and add a column containing the sheetnames

precipitation_2018_2019 <- precipitation_2018_2019 %>% 
  na.omit(Total) %>% #omit rows without precipitation data
  mutate(Year = substr(sheetname, 1, 4), Month = month.name[Month]) %>%  #add a variable for year and convert month to a character variable
  select(Year, Month, Total) %>% 
  as_tibble()
```

There are `r nrow(mr_trash_wheel_df)` observations in the Mr. Trash Wheel dataset and `r nrow(precipitation_2018_2019)` observations in the precipitation dataset for year 2018 and 2019. 

Some of the key variables in the Mr. Trash Wheel dataset are: \
`year`  and `month` represents the year and month of trash collection\
`dumpster` column includes specific dumpster number \
`sports_balls` column includes the number of sports balls collected in Mr. Trash Wheel on the given specific date and dumpster. The mean number of sports balls is `r mean(pull(filter(mr_trash_wheel_df), sports_balls))`. And the median number of sports balls in the dumpster in 2019 specifically is `r median (pull(filter(mr_trash_wheel_df, year == 2019), sports_balls))`\
`weight_tons` represents the weight of trash in tons collected by Mr. Trash Wheel on a given dumpster and date \

Some of the key variables in the precipitation dataset for year 2018 and 2019 are:\
`Year` represents the year of precipitation measuremens\
`Month` is the month of precipitation measurement\
`Total` column includes the total precipitation for the year-month\

The total precipitation measured in 2018 is `r sum(pull(filter(precipitation_2018_2019, Year == 2018), Total))` 

\newpage
# Problem 2 
## read and clean pols_month
```{r read and clean pols_month, message=F, warning=F}
rm(list = ls()) #clean environment to start a new problem

pols_month <- read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% #load pols_month dataset
  janitor::clean_names() 

pols_month <- pols_month %>%
  separate(mon, sep = "-", into = c("year", "month","day")) %>% #Use separate() to break up the variable mon into integer variables year, month, and day
  mutate(
    year = as.integer(year), 
    month = as.integer(month), 
    day = as.integer(day),
    month = month.name[month]) %>% 
  pivot_longer(c("prez_gop","prez_dem"), #a president variable taking values gop and dem
    names_to = "president",
    values_to = "president_boolean",
    names_prefix = "prez_"
  ) %>% 
  filter(president_boolean != 0) %>% 
  select(-c(president_boolean, day)) #remove the day variable
```

## read and clean the data in snp.csv
```{r read and clean the data in snp.csv, message=F, warning=F}
snp <- read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% #load snp dataset
  janitor::clean_names() 

snp <- snp %>%
  mutate(
    year = lubridate::year(lubridate::mdy(date)), 
    month = lubridate::month(lubridate::mdy(date)), 
    month = month.name[month]) %>% 
  select(year, month, close) %>% #year and month are the leading columns
  arrange(year, month) #arrange according to year and month
```

## read and tidy unemplotment data
```{r read and tidy unemplotment data, message=F,warning=F}
unemployment <- read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() 

unemployment <- unemployment %>% 
  pivot_longer(
    jan:dec, 
    names_to = 'month', 
    values_to = "unemployment_rate"
  ) %>% 
  mutate(
    year = as.integer(year),
    month = apply(as.matrix(month), 1, function(abb){
      s <- str_to_title(abb)
      for (i in 1:12) {
        if (month.abb[i] == s ) {
          return(month.name[i])
        }
      }
    }
    ))

```

## Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r, message=F,warning=F}
fivethirtyeight_merged <- pols_month %>% 
  left_join(snp, by = c("year", "month")) %>% 
  left_join(unemployment, by = c("year", "month"))
```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

\ The pols-month dataset contains `r nrow(pols_month)` observations of `r ncol(pols_month)` variables related to the number of national politicians who are democratic or republican at the given year-month. The newly created variable `president` in this dataset indicates whether the president in office in the associated year-month was republican (`gop`) or democratic (`dem`)

\ The snp dataset contains `r nrow(snp)` observations of `r ncol(snp)` variables related to Standard & Poor’s stock market index on the given year-month.

\ The unemployment dataset contains `r nrow(unemployment)` observations of `r ncol(unemployment)` variables including the percentage of unemployment in a given month and year.

\ All these 3 datasets were merged to a final dataframe called `fivethirtyeight_merged`. This final data frame contains `r nrow(fivethirtyeight_merged)` observations of `r ncol(fivethirtyeight_merged)` variables. The data years range from year `r min(pull(fivethirtyeight_merged, year))` to year `r max(pull(fivethirtyeight_merged, year))`. Overall, this final data frame includes the number of senator, representative, governor of each party, party of the president in office, the closing values of the S&P stock index, and the unemployment rate on the associated year and month. Key variables: 

* `close` = the closing values of the S&P stock index on the associated month and year
* `unemplyment_rate` = percentage of unemployment of the associated month and year
* `president` = whether the president in office in the associated year-month was republican (`gop`) or democratic (`dem`)
* `gov_gop`: the number of republican governors on the associated month and year
* `sen_gop`: the number of republican senators on the associated month and year
* `rep_gop`: the number of republican representatives on the associated month and year
* `gov_dem`: the number of democratic governors on the associated month and year
* `sen_dem`: the number of democratic senators on the associated month and year
* `rep_dem`: the number of democratic representatives on the associated month and year

# Problem 3
## Load and tidy popular_baby_names data.

```{r Load and tidy popular_baby_names data, message=F, warning=F}
popular_baby_names <- read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() 

table(popular_baby_names$gender) #gender column seems fine

table(popular_baby_names$ethnicity) #needs to be cleaned 

popular_baby_names <- popular_baby_names %>%
    mutate(
    childs_first_name = str_to_title(childs_first_name),
    ethnicity = recode(ethnicity,
  "BLACK NON HISP" = "BLACK NON HISPANIC", 
  "WHITE NON HISP" = "WHITE NON HISPANIC", 
  "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER")) %>%
  distinct(year_of_birth, rank, gender, childs_first_name, ethnicity) #ensures that each row is unique

table(popular_baby_names$ethnicity) #cleaned

```


